### 2021-07-01 doc/src/MSLevel.tex
Tweak *topi CSRs, to simplify and allow IPRIO = 0
Modified field IPRIO of CSRs mtopi, stopi, vstopi in two ways, for the
case when IPRIO is more informative than just always 0 or 1:  First,
large "priority numbers" (indicating low priority) are always saturated
to 255 instead of implementation-specific limits MMAXIPRIO and
SMAXIPRIO.  Second, the highest-priority interrupts with priority
number 0 now report IPRIO = 0, distinct from the next-lower priority
number 1.

Modified CSR hvicontrol correspondingly.  Also added field IPRIOM
(IPRIO Mode) to hvicontrol to support reporting vstopi.IPRIO al

The value of field IPRIO depends on parameter \emph{MMAXIPRIO}.
If not all bytes of the machine-level \z{iprio} array are read-only
zeros, MMAXIPRIO is the maximum allowed value of an \z{iprio} byte
(=~${\mbox{2}^\textrm{IPRIOLEN}-\mbox{1}}$).
On the other hand, if all bytes of the machine-level \z{iprio} array
are read-only zeros, an implementation may define MMAXIPRIO either as
${\mbox{2}^\textrm{IPRIOLEN}-\mbox{1}}$ or as~1.

When \z{mtopi} is not zero, if the priority number for the reported
interrupt is in the range 1 to MMAXIPRIO, IPRIO is simply that number.
If the interrupt's priority number is zero or greater than MMAXIPRIO,
IPRIO is saturated to either 1 or MMAXIPRIO as follows:
\begin{itemize}

\item
If the interrupt's priority number is greater than MMAXIPRIO, then
IPRIO is MMAXIPRIO (lowest priority).

\item
If the interrupt's priority number is zero and interrupt number IID has
a default priority higher than a machine external interrupt, then IPRIO
is~1 (highest priority).

\item
If the interrupt's priority number is zero and interrupt number IID has
a default priority lower than a machine external interrupt, then IPRIO
is MMAXIPRIO (lowest priority).

@@

Note that if MMAXIPRIO =~1, IPRIO is always~1 whenever \z{mtopi} is not
zero.

@@

The value of field IPRIO depends on parameter \emph{SMAXIPRIO}.
If not all bytes of the supervisor-level \z{iprio} array are read-only
zeros, SMAXIPRIO is the maximum allowed value of an \z{iprio} byte
(=~${\mbox{2}^\textrm{IPRIOLEN}-\mbox{1}}$).
On the other hand, if all bytes of the supervisor-level \z{iprio} array
are read-only zeros, an implementation may define SMAXIPRIO either as
${\mbox{2}^\textrm{IPRIOLEN}-\mbox{1}}$ or as~1.

When \z{stopi} is not zero, if the priority number for the reported
interrupt is in the range 1 to SMAXIPRIO, IPRIO is simply that number.
If the interrupt's priority number is zero or greater than SMAXIPRIO,
IPRIO is saturated to either 1 or SMAXIPRIO as follows:
\begin{itemize}

\item
If the interrupt's priority number is greater than SMAXIPRIO, then
IPRIO is SMAXIPRIO (lowest priority).

\item
If the interrupt's priority number is zero and interrupt number IID has
a default priority higher than a supervisor external interrupt, then
IPRIO is~1 (highest priority).

\item
If the interrupt's priority number is zero and interrupt number IID
has a default priority lower than a supervisor external interrupt, then
IPRIO is SMAXIPRIO (lowest priority).

@@

Note that if SMAXIPRIO =~1, IPRIO is always~1 whenever \z{stopi} is not
zero.

>>>>

If all bytes of the machine-level \z{iprio} array are read-only zeros,
a simplified implementation of field IPRIO is allowed in which its
value is always 1 whenever \z{mtopi} is not zero.

Otherwise, when \z{mtopi} is not zero, if the priority number for the reported
interrupt is in the range 1 to 255, IPRIO is simply that number.
If the interrupt's priority number is zero or greater than 255,
IPRIO is set to either 0 or 255 as follows:
\begin{itemize}

\item
If the interrupt's priority number is greater than 255, then
IPRIO is 255 (lowest priority).

\item
If the interrupt's priority number is zero and interrupt number IID has
a default priority higher than a machine external interrupt, then IPRIO
is~0 (highest priority).

\item
If the interrupt's priority number is zero and interrupt number IID has
a default priority lower than a machine external interrupt, then IPRIO
is 255 (lowest priority).

@@

@@

If all bytes of the supervisor-level \z{iprio} array are read-only
zeros, a simplified implementation of field IPRIO is allowed in which
its value is always 1 whenever \z{stopi} is not zero.

Otherwise, when \z{stopi} is not zero, if the priority number for the reported
interrupt is in the range 1 to 255, IPRIO is simply that number.
If the interrupt's priority number is zero or greater than 255,
IPRIO is set to either 0 or 255 as follows:
\begin{itemize}

\item
If the interrupt's priority number is greater than 255, then
IPRIO is 255 (lowest priority).

\item
If the interrupt's priority number is zero and interrupt number IID has
a default priority higher than a supervisor external interrupt, then
IPRIO is~0 (highest priority).

\item
If the interrupt's priority number is zero and interrupt number IID
has a default priority lower than a supervisor external interrupt, then
IPRIO is 255 (lowest priority).

@@

### 2021-07-01 doc/src/VSLevel.tex
Tweak *topi CSRs, to simplify and allow IPRIO = 0
Modified field IPRIO of CSRs mtopi, stopi, vstopi in two ways, for the
case when IPRIO is more informative than just always 0 or 1:  First,
large "priority numbers" (indicating low priority) are always saturated
to 255 instead of implementation-specific limits MMAXIPRIO and
SMAXIPRIO.  Second, the highest-priority interrupts with priority
number 0 now report IPRIO = 0, distinct from the next-lower priority
number 1.

Modified CSR hvicontrol correspondingly.  Also added field IPRIOM
(IPRIO Mode) to hvicontrol to support reporting vstopi.IPRIO al

The format of \z{hvicontrol} is:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
bit 30     & VTI \\
bits 27:16 & IID (\WARL) \\
bits 7:0   & IPRIO \\
\end{displayLinesTable}
All other bits of \z{hvicontrol} are reserved and read as zeros.

@@

All \z{hvicontrol} fields, VTI, IID, and IPRIO, together can affect
the value of CSR \z{vstopi} (Virtual Supervisor Top Interrupt) and
therefore the interrupt identity reported in \z{vscause} when an
interrupt traps to \mbox{VS-mode}.
IID is a {\WARL} unsigned integer field with at least 6 implemented
bits, while IPRIO is always the full 8~bits.

@@

\item
if \z{hvicontrol} fields VTI =~1, $\mbox{IID} \neq \mbox{9}$, and
$\mbox{IPRIO} \neq \mbox{0}$:
\begin{tightList}
\item[+]
the major interrupt specified by \z{hvicontrol} fields IID and IPRIO.
\end{tightList}

@@

When no interrupt candidates satisfy the conditions of the list above,
\z{vstopi} is zero.
Else, \z{vstopi} fields IID and IPRIO are determined by the
highest-priority interrupt from among the candidates.
The usual priority order for supervisor level applies, as specified by
Table~\ref{tab:intrPrios-S} on page~\pageref{tab:intrPrios-S}, except
that priority numbers are taken from the candidate list above, not from
the supervisor-level \z{iprio} array.
As always, ties in nominal priority are broken by the default priority
order from Section~\ref{sec:majorIntrs}.
If the priority number for the highest-priority candidate is within the
range 1 to 255, IPRIO in \z{vstopi} is that value;
else, IPRIO is saturated to either 1 or 255 in the manner documented
for \z{stopi} in Section~\ref{sec:stopi}, assuming parameter
SMAXIPRIO = 255.

>>>>

The format of \z{hvicontrol} is:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
bit 30     & VTI \\
bits 27:16 & IID (\WARL) \\
bit 8      & IPRIOM \\
bits 7:0   & IPRIO \\
\end{displayLinesTable}
All other bits of \z{hvicontrol} are reserved and read as zeros.

@@

All \z{hvicontrol} fields together can affect
the value of CSR \z{vstopi} (Virtual Supervisor Top Interrupt) and
therefore the interrupt identity reported in \z{vscause} when an
interrupt traps to \mbox{VS-mode}.
IID is a {\WARL} unsigned integer field with at least 6 implemented
bits, while IPRIO is always the full 8~bits.

@@

\item
if \z{hvicontrol} fields VTI =~1 and $\mbox{IID} \neq \mbox{9}$:
\begin{tightList}
\item[+]
the major interrupt specified by \z{hvicontrol} fields IID and IPRIO.
\end{tightList}

@@

\begin{commentary}
When both \z{hstatus}.VGEIN =~0 and \z{hvicontrol}.VTI =~1, the
absence of an interrupt for VS level can be indicated only by setting
\z{hvicontrol} fields IID =~9 and IPRIO~=~0.
Software might want to use the pair IID =~9, IPRIO~=~0 generally to
represent\/ \emph{no interrupt} in \z{hvicontrol}.
\end{commentary}

When no interrupt candidates satisfy the conditions of the list above,
\z{vstopi} is zero.
Else, \z{vstopi} fields IID and IPRIO are determined by the
highest-priority interrupt from among the candidates.
The usual priority order for supervisor level applies, as specified by
Table~\ref{tab:intrPrios-S} on page~\pageref{tab:intrPrios-S}, except
that priority numbers are taken from the candidate list above, not from
the supervisor-level \z{iprio} array.
As always, ties in nominal priority are broken by the default priority
order from Section~\ref{sec:majorIntrs}.
If bit IPRIOM (IPRIO Mode) of \z{hvicontrol} is zero,
IPRIO in \z{vstopi} is~1;
else, if the priority number for the highest-priority candidate is within the
range 1 to 255, IPRIO is that value;
else, IPRIO is set to either 0 or 255 in the manner documented
for \z{stopi} in Section~\ref{sec:stopi}.
